### 2022-01-26 cv/cv.tex
Add CV template

\documentclass[10pt, a4paper, oneside]{article}

\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=black, pdfauthor=Evgenii Iuliugin,
  pdftitle=Evgenii Iuliugin's CV}

\usepackage{indentfirst}
\usepackage[top=1.0cm, bottom=1.0cm, left=1.0cm, right=1.0cm,
  footskip=1.0cm]{geometry}

\usepackage{titlesec}
\titleformat
    {\section} % command
    [display] % shape
    {\normalfont\bfseries} % format
    {} % label
    {} % separator
    {\vspace{-2ex}} % before-code
    [
      \vspace{-2ex}
      \rule{\textwidth}{0.3pt}
    ] % after-code

\setlength\parindent{0pt}

\pagestyle{empty}

\begin{document}

\Huge
Evgenii Iuliugin

\LARGE
\textit{Software Engineer}

\begin{flushright}
  \large
  \href{mailto:yulyugin@gmail.com}{yulyugin@gmail.com}

  % phone number
\end{flushright}

\section*{SUMMARY}

\section*{EXPERIENCE}

\section*{EDUCATION}

\section*{PUBLICATIONS}

\section*{OTHER}

\end{document}

### 2022-02-08 cv/cv.tex
Fill in CV parts

@@
-\documentclass[10pt, a4paper, oneside]{article}
+\documentclass[12pt, a4paper, oneside]{article}
 
@@
 
+\usepackage{tabularx}
+\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
+
+\usepackage{enumitem}
+\setlist[itemize]{noitemsep, topsep=-1em, leftmargin=2.5\labelsep}
+
 \begin{document}
@@
 \Huge
-Evgenii Iuliugin
+\textbf{Evgenii Iuliugin}
 
-\LARGE
+\Large
 \textit{Software Engineer}
@@
 
+\normalsize
+
 \section*{SUMMARY}
+
+Software engineer with deep knowledge of hardware modeling and virtualization,
+operating system internals, hardware architectures, and software design
+principles. Result-oriented and well-qualified to develop complicated software
+products from gathering technical requirements to full system integration
+and debug.
+
+Programming languages: C/C++ (10 years of experience); x86 assembly; Python.
+
+Areas of interest: virtualization, operating systems, compiler technologies.
 
@@
 
+\begin{table}[htpb]
+  \begin{tabularx}{\textwidth}{lY}
+    2012 --- 2016 & \textbf{Intel Corporation} \hfill Moscow, Russia \\
+    & \textbf{\textit{Software Engineer}} \\
+    2016 --- now  & \textbf{Intel Corporation} \hfill Stockholm, Sweden \\
+    & \textbf{\textit{Senior Software Engineer}} \\
+    & \textit{Designed, implemented and optimized processor models. Integrated
+    the models with virtual platforms used for pre-silicon validation.} \\
+    & \begin{itemize}
+      \item Led development of VT-x based acceleration in Intel Simics.
+      \item Designed and implemented VT-x based acceleration in Intel Simics
+        for macOS.
+      \item Designed and implemeneted automatic generation of x86 instruction
+        descriptions for Intel Simics using Intel XED instruction descriptions.
+      \item Analyzed and improved performance of various Intel Simics processor
+        and device models.
+      \item Debugged and enabled boot scenario for different operating systems
+        and hypervisors on Intel Simics.
+      \item Collected and analyzed customer requirements, designed model
+        components based on the requirements.
+      \item Consulted team members on operating system internals.
+      \item Regularly shared knowledge via technical presentations to the team.
+      \item Regularly reviewed designs and code from peers.
+      \item Mentored junior contributors.
+      \end{itemize} \\
+  \end{tabularx}
+\end{table}
+
 \section*{EDUCATION}
+
+\begin{table}[htpb]
+  \begin{tabularx}{\textwidth}{lY}
+    2009 --- 2015 & \textbf{Moscow Institute of Physics and Technology} \hfill Moscow, Russia \\
+    & \textit{MS in Applied Mathematics and Physics} \\
+    2015 --- 2019 & \textbf{Moscow Institute of Physics and Technology} \hfill Moscow, Russia \\
+    & \textit{PhD in Computer Science} \\
+  \end{tabularx}
+\end{table}
+
+\section*{LANGUAGES}
+
+\begin{itemize}
+\item Russian --- fluent;
+\item English --- intermediate;
+\item Swedish --- basic.
+\end{itemize}
+
+\section*{SKILLS}
+
+\begin{itemize}
+\item Expert knowledge of virtual machine design and implementation,
+  including binary translation and direct execution.
+\item Knowledge of operating system design and implementation.
+\item Deep knowledge of IA-32 and AMD64 processor architectures.
+\item Cross-platform development for Windows, Linux and macOS with both native
+  APIs and platform-agnostic libraries.
+\item Advanced experience with driver development, including Windows Driver
+  Model (WDM).
+\item Collaborative development with Subversion, GIT, Jenkins, TeamCity, etc.
+\item Proficiency with configuration management and build systems based on MAKE.
+\end{itemize}
 
@@
 
-\section*{OTHER}
-
 \end{document}

### 2023-01-01 _posts/papers-2022.tex
Publish 2022 paper summary

\documentclass[12pt, a4paper, oneside]{article}

\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=blue, pdfauthor=Evgenii Iuliugin}

\usepackage{indentfirst}
\usepackage[top=2.0cm, bottom=2.0cm, left=2.0cm, right=2.0cm,
  footskip=2.0cm]{geometry}

\pagestyle{empty}

\usepackage[bibencoding=inputenc, backend=biber, language=auto, style=verbose]{biblatex}
\newcommand{\bibpath}{../bibliography}
\addbibresource{\bibpath/Machine Learning/hardware.bib}
\addbibresource{\bibpath/Machine Learning/optimizations.bib}
\addbibresource{\bibpath/Processors/riscv.bib}
\addbibresource{\bibpath/Simulators/cycle-accurate.bib}
\addbibresource{\bibpath/Simulators/description-languages.bib}
\addbibresource{\bibpath/Simulators/functional.bib}
\addbibresource{\bibpath/Simulators/frameworks.bib}
\addbibresource{\bibpath/Simulators/target-analysis.bib}
\addbibresource{\bibpath/compilers.bib}
\addbibresource{\bibpath/embedded-systems.bib}
\addbibresource{\bibpath/fun.bib}
\addbibresource{\bibpath/performance.bib}
\addbibresource{\bibpath/software-philosophy.bib}
\addbibresource{\bibpath/virtual-machines.bib}

\begin{document}

\section*{Machine Learning}
\begin{itemize}
    \item A new reconfigurable hardware architecture designed specifically to facilitate machine learning requirements:

    \cite{Prabhakar:Plasticine:2017}.

    \item An optimization to increase parallelism of machine learning tasks:

    \cite{Yuanzhong:Cross-Replica-Sharding:2020}.
\end{itemize}

\section*{Processors}
\begin{itemize}
    \item Western Digital builds its own RISC-V processor:

    \cite{Wheeler:WD-RISC-V:2019}.
\end{itemize}

\section*{Simulators}
\begin{itemize}
    \item Basic hardware performance modeling approach with separation of functional and timing simulation using ports:

    \cite{Emer:Asim:2002}.

    \item Micro architectural processor simulation framework:

    \cite{Hughes:Rsim:2002}.

    \item Infrastructure and framework for cycle-accurate and functional simulation:

    \cite{Austin:SimpleScalar:2002}.

    \item KVM-based direct execution for ARM:

    \cite{Junger:ARM-on-ARM:2020}.

    \item Overview of Simics simulator:

    \cite{Magnusson:Simics:2002}.

    \item Basic algorithm for multiprocessor target on multiprocessor host simulation:

    \cite{Kenney:MP-on-MP:2015}.

    \item Single instruction JIT compilation approach:

    \cite{Nohl:JIT-Simulation:2002}.

    \item Memory and thread usage analysis for software running inside a virtual platform:

    \cite{Dahl:VP-Sanitizer:2018}.
\end{itemize}

\section*{Compilers}
\begin{itemize}
    \item An amazing overview of basic but demonstrative optimizations performed by a modern C/C++ compiler:

    \cite{Godbolt:Optimizations:2020}.
\end{itemize}

\section*{Embedded Systems}
\begin{itemize}
    \item PhD thesis by Jakob Engblom focused on studying of worst case execution time analysis for embedded systems:

    \cite{Engblom:WCET:2002}.
\end{itemize}

\section*{Fun}
\begin{itemize}
    \item Really fun connection between texts of several popular songs and computer basics:

    \cite{Colwell:Rock-N-Roll:2002}.
\end{itemize}

\section*{Performance}
\begin{itemize}
    \item A way to determine maximum performance of a particular system together with few ideas on how to decide what optimizations make sense to do and when:

    \cite{Williams:Roofline:2009}.
\end{itemize}

\section*{Software Philosophy}
\begin{itemize}
    \item Discussions on what a perfect design process show look like:

    \cite{Parnas:Fake-It:1986}.

    \item Philosophical discussions about causes of errors in software:

    \cite{Horner:Error-Rates:2019}.
\end{itemize}

\section*{Virtual Machines}
\begin{itemize}
    \item A set of techniques to detect a virtualized Android environment using basic phone identification information, sensor data and timing of processor instruction execution:

    \cite{Petsas:Android-VM-Detection:2014}.
\end{itemize}

\end{document}

### 2023-01-03 _posts/papers-2023.tex
Add short description for ericsson VP paper

\input{papers.tex}

\begin{document}

\section*{Simulators}

\begin{itemize}
    \item Ericsson's experience with virtual platform for pre-silicon software development. The virtual platforms are based on SystemC/TLM. This paper may be used as a showcase of simulation importance.

    \cite{Dahl:Ericsson-VP:2016}.
\end{itemize}

\end{document}

### 2023-01-04 _posts/papers-2023.tex
Add ARM Neoverse citation

@@
 The virtual platforms are based on SystemC/TLM.
-This paper may be used as a showcase of simulation importance.
+This paper may be used as a showcase of simulation importance:
 
@@
 
+\section*{Processors}
+\begin{itemize}
+    \item A 64 core ARM reference server CPU design similar to Cortex-A76 microarchitecture.
+Single core performance is not that good but the processor is expectedly much more power efficient.
+New hardware-based coherence model for i-cache aiming improvement for multiple VM performance.
+SPEC CPU 2006 is still referenced as a benchmark.
+Interesting reference to read --- <<Editorial: Intel Hides Behind ICC>> --- MPR claims that ICC must be approximately 30\% faster than GCC (at least for SPEC workloads):
+
+    \cite{Gwennap:ARM-Neoverse:2019}.
+\end{itemize}
+
 \end{document}

### 2023-01-07 _posts/papers-2023.tex
Add KVM/ARM to summary of 2023 papers

@@
     \cite{Dahl:Ericsson-VP:2016}.
+
+    \item The paper describes an early design of KVM?ARM port using the first version of ARM virtualization hardware support.
+Authors claim to introduce a novel "split-mode" virtualization approach as a way to run hypervisor simultaneously in different CPU privileged modes.
+This approach is not new though in my opinion.
+There exists for a long time hypervisors like Simics that use both kernel and user CPU modes to leverage different parts of CPU virtualization.
+Nevertheless, authors have done a grade work adding ARM support to KVM with generic and rather minimal changes to KVM and Linux kernel.
+There is an interesting comparison of x86 and ARM virtualization power and performance overheads in the paper.
+Some of the links in references are also worse looking through later.
+
+    \cite{Dall:KVM-ARM:2014}
 \end{itemize}

### 2023-01-13 _posts/papers-2023.tex
Add ispc citation

@@
 
+\section*{Compilers}
+\begin{itemize}
+    \item The paper presents ispc --- Intel SPMD Program Compiler.
+The paper introduces new SPMD-on-SIMD concept that seems to be quite reasonable for running effectively compiling SPMD programs for CPUs.
+A few specific optimizations are introduced to enable efficient usage of a CPU's SIMD hardware unit.
+The project looks like an interesting alternative to hand-written intrinsic-based optimizations.
+The project is still alive at the moment of writing the note.
+I can also see a few well familiar names in the list of contributors.
+As a side note, the paper can be a good reference for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course>>} if that one ever gets back to live:
+
+    \cite{Pharr:ispc:2012}.
+\end{itemize}
+
 \end{document}

### 2023-01-14 _posts/papers-2023.tex
Fix quotes

@@
 I can also see a few well familiar names in the list of contributors.
-As a side note, the paper can be a good reference for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course>>} if that one ever gets back to live:
+As a side note, the paper can be a good reference for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course} if that one ever gets back to live:
 

### 2023-01-14 paper-reviews/papers-2023.tex
Improve ispc description

--- 2023-01-07 _posts/papers-2023.tex
@@
 
+\section*{Compilers}
+\begin{itemize}
+    \item The paper presents ispc --- Intel SPMD Program Compiler.
+A new <<SPMD-on-SIMD>> concept is introduced.
+The concept seems to be quite reasonable for efficient compilation of SPMD programs for CPUs.
+A few specific optimizations are described to enable better usage of a CPU's SIMD hardware unit.
+The project looks like an interesting alternative to hand-written intrinsic-based optimizations.
+The project is still alive at the moment of writing the note.
+I can also see a few well familiar names in the list of contributors.
+As a side note for myself, the paper can be a good reference for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course} if that one ever gets back to live:
+
+    \cite{Pharr:ispc:2012}.
+\end{itemize}
+
 \end{document}

### 2023-01-18 paper-reviews/papers-2023.tex
Count total number of citations
Add an article about test case reduction in C

@@
 \begin{document}
+
+A total of \total{citenum} papers read in the year.
 
@@
     \cite{Pharr:ispc:2012}.
+
+    \item As practice shows creating a concise reproducer for a bug is sometimes as complex as fixing a bug.
+Compiler bugs can be especially hard to report as they are usually observed on real and rather big projects.
+Even relatively small \href{https://yulyugin.github.io/blog/2022/icc-bug}{compiler bug} reproducers could be hard to analyze and report.
+Authors of the paper tried to address this important issue by implementing an automatic test case reducers for C:
+
+    \cite{Regehr:C-Reduce:2012}
 \end{itemize}

### 2023-01-18 paper-reviews/papers-2023.tex
Add ARM Helios article

@@
     \cite{Gwennap:ARM-Neoverse:2019}.
+
+    \item The article describes ARM 64-bit only architecture called Helios.
+It is a first licensable CPU that combines OOO and multithreaded execution.
+The main application for the processor are data transfer infrastructure and automotive industry.
+The processor architecture supports INT8 and FP16 formats that are the current trend for DNN acceleration.
+On top of INT8 and FP16 the processor can be closes integrated with various accelerators:
+
+    \cite{Demler:Helios:2019}
 \end{itemize}

### 2023-01-26 paper-reviews/papers-2023.tex
Add Philips VLIW report
Add multiflow VLIW paper

@@
 
-    \cite{Dall:KVM-ARM:2014}
+    \cite{Dall:KVM-ARM:2014}.
 \end{itemize}
@@
 
-    \cite{Demler:Helios:2019}
+    \cite{Demler:Helios:2019}.
 \end{itemize}
@@
 
-    \cite{Regehr:C-Reduce:2012}
+    \cite{Regehr:C-Reduce:2012}.
+\end{itemize}
+
+\section*{Computer Architecture}
+\begin{itemize}
+    \item Nice high level overview of VLIW architecture and comparison with typical RISC and CISC designs.
+Great reference for students to show the basic concepts of CPU architectures even though it is not fully relevant anymore:
+
+    \cite{Philips:VLIW:1997}.
+
+    \item The paper describes a VLIW architecture developed by Multiflow Corporation.
+It is one of the VLIW pioneering designs.
+It is always enjoyable to read old papers about computer architecture trends and ideas; compare the old ideas and expectations with how it turned out now: <<a VLIW machine was expected to provide from ten to thirty times the performance of a more conventional machine built of the same implementation technology>>.
+A great comments for the citation would be something like: <<if only we would have smart enough compiler>>.
+Industry practice so far showed that VLIW could be an amazing architecture for some DSPs while a superscalar processor delivers better results for general purpose usage:
+
+    \cite{Colwell:Multiflow-VLIW:1987}.
 \end{itemize}

### 2023-01-27 paper-reviews/papers-2023.tex
Add mobileye article

@@
 
+\section*{Autonomous Driving}
+\begin{itemize}
+    \item The article present an overview of Mobileye existing and in flight processors for advanced-driver-assistance-systems (ADAS).
+The article also presents a simplistic comparison with Nvidia GPU-based solutions for the area.
+The main comparison metrics are trillion operations per second (TOPS) and efficiency which is measured as TOPS divided by power consumption.
+It is an overall nice basic overview of what is happening in the area with cool examples of 360-degree image construction and object classification:
+
+    \cite{Delmer:Mobileye:2019}.
+\end{itemize}
+
 \end{document}

### 2023-01-29 paper-reviews/papers-2023.tex
Add optical AI paper

@@
 
-    \item As practice shows creating a concise reproducer for a bug is sometimes as complex as fixing a bug.
+    \item As practice shows creating a concise reproducer for a bug is sometimes as complex as fixing the bug.
 Compiler bugs can be especially hard to report as they are usually observed on real and rather big projects.
@@
 
+\section*{Machine Learning}
+\begin{itemize}
+    \item Optical computing for AI moves closer to commercialization.
+Startups originating from MIT and Princeton demonstrate interesting optical computing for AI solutions.
+The systems are targeting convolutional neural networks.
+2D filtering is equal to 2D convolution.
+The approach is aimed to reduce heat and power consumption and of course improve performance simultaneously.
+One of the main problems so far is optical to electrical and electrical to optical signal conversion.
+Solutions for the problem seems to be a key for high performance at the moment.
+The article also contains a link to a scientific paper with more details.
+
+    \cite{Wheeler:Optical-AI:2019}.
+\end{itemize}
+
 \end{document}

### 2023-02-01 paper-reviews/papers-2023.tex
Add Centaur AI accelerator paper

@@
     \cite{Wheeler:Optical-AI:2019}.
+
+    \item Centaur has developed an x86 processor with deep-learning accelerator --- Ncore.
+The accelerator has a typical design for the field of application: it's a VLIW architecture relying on traditional SIMD approach with smaller sized vector elements like INT8, BF16, etc.
+The architecture is 64 times wider that Intel's AVX-512.
+The x86 cores are alike low-end Skylake cores but it can achieve better neural network performance thanks to the accelerator.
+Centaur has also constructed a basic software stack that convert TensorFlow to an internal graph format that is later compiled into Ncore assembly - another typical approach for deep-learning accelerators.
+The article reports that so fat Centaur's software can only handle inference but not training.
+
+    \cite{Gwennap:Centaur-AI:2019}.
 \end{itemize}

### 2023-02-01 paper-reviews/papers-2023.tex
Reformat paper reviews

@@
 \begin{itemize}
-    \item Ericsson's experience with virtual platform for pre-silicon software development.
-The virtual platforms are based on SystemC/TLM.
-This paper may be used as a showcase of simulation importance:
+    \item \cite{Dahl:Ericsson-VP:2016}:
 
-    \cite{Dahl:Ericsson-VP:2016}.
+    Ericsson's experience with virtual platform for pre-silicon software development. The virtual platforms are based on SystemC/TLM. This paper may be used as a showcase of simulation importance.
 
-    \item The paper describes an early design of KVM?ARM port using the first version of ARM virtualization hardware support.
+    \item \cite{Dall:KVM-ARM:2014}:
+
+    The paper describes an early design of KVM?ARM port using the first version of ARM virtualization hardware support.
 Authors claim to introduce a novel "split-mode" virtualization approach as a way to run hypervisor simultaneously in different CPU privileged modes.
@@
 Some of the links in references are also worse looking through later.
-
-    \cite{Dall:KVM-ARM:2014}.
 \end{itemize}
@@
 \begin{itemize}
-    \item A 64 core ARM reference server CPU design similar to Cortex-A76 microarchitecture.
+    \item \cite{Gwennap:ARM-Neoverse:2019}:
+
+    A 64 core ARM reference server CPU design similar to Cortex-A76 microarchitecture.
 Single core performance is not that good but the processor is expectedly much more power efficient.
@@
 SPEC CPU 2006 is still referenced as a benchmark.
-Interesting reference to read --- <<Editorial: Intel Hides Behind ICC>> --- MPR claims that ICC must be approximately 30\% faster than GCC (at least for SPEC workloads):
+Interesting reference to read --- <<Editorial: Intel Hides Behind ICC>> --- MPR claims that ICC must be approximately 30\% faster than GCC (at least for SPEC workloads).
 
-    \cite{Gwennap:ARM-Neoverse:2019}.
+    \item \cite{Demler:Helios:2019}:
 
-    \item The article describes ARM 64-bit only architecture called Helios.
+    The article describes ARM 64-bit only architecture called Helios.
 It is a first licensable CPU that combines OOO and multithreaded execution.
@@
 The processor architecture supports INT8 and FP16 formats that are the current trend for DNN acceleration.
-On top of INT8 and FP16 the processor can be closes integrated with various accelerators:
+On top of INT8 and FP16 the processor can be closes integrated with various accelerators.
 
-    \cite{Demler:Helios:2019}.
 \end{itemize}
@@
 \begin{itemize}
-    \item The paper presents ispc --- Intel SPMD Program Compiler.
+    \item \cite{Pharr:ispc:2012}:
+
+    The paper presents ispc --- Intel SPMD Program Compiler.
 A new <<SPMD-on-SIMD>> concept is introduced.
@@
 I can also see a few well familiar names in the list of contributors.
-As a side note for myself, the paper can be a good reference for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course} if that one ever gets back to live:
+As a side note for myself, the paper can be a good reference for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course} if that one ever gets back to live.
+\end{itemize}
 
-    \cite{Pharr:ispc:2012}.
+\section*{Testing}
+\begin{itemize}
+    \item \cite{Regehr:C-Reduce:2012}:
 
-    \item As practice shows creating a concise reproducer for a bug is sometimes as complex as fixing the bug.
+    As practice shows creating a concise reproducer for a bug is sometimes as complex as fixing the bug.
 Compiler bugs can be especially hard to report as they are usually observed on real and rather big projects.
 Even relatively small \href{https://yulyugin.github.io/blog/2022/icc-bug}{compiler bug} reproducers could be hard to analyze and report.
-Authors of the paper tried to address this important issue by implementing an automatic test case reducers for C:
-
-    \cite{Regehr:C-Reduce:2012}.
+Authors of the paper tried to address this important issue by implementing an automatic test case reducers for C.
 \end{itemize}
@@
 \begin{itemize}
-    \item Nice high level overview of VLIW architecture and comparison with typical RISC and CISC designs.
-Great reference for students to show the basic concepts of CPU architectures even though it is not fully relevant anymore:
+    \item \cite{Philips:VLIW:1997}:
 
-    \cite{Philips:VLIW:1997}.
+    Nice high level overview of VLIW architecture and comparison with typical RISC and CISC designs.
+Great reference for students to show the basic concepts of CPU architectures even though it is not fully relevant anymore.
 
-    \item The paper describes a VLIW architecture developed by Multiflow Corporation.
+    \item \cite{Colwell:Multiflow-VLIW:1987}:
+
+    The paper describes a VLIW architecture developed by Multiflow Corporation.
 It is one of the VLIW pioneering designs.
@@
 A great comments for the citation would be something like: <<if only we would have smart enough compiler>>.
-Industry practice so far showed that VLIW could be an amazing architecture for some DSPs while a superscalar processor delivers better results for general purpose usage:
-
-    \cite{Colwell:Multiflow-VLIW:1987}.
+Industry practice so far showed that VLIW could be an amazing architecture for some DSPs while a superscalar processor delivers better results for general purpose usage.
 \end{itemize}
@@
 \begin{itemize}
-    \item The article present an overview of Mobileye existing and in flight processors for advanced-driver-assistance-systems (ADAS).
+    \item \cite{Delmer:Mobileye:2019}:
+
+    The article present an overview of Mobileye existing and in flight processors for advanced-driver-assistance-systems (ADAS).
 The article also presents a simplistic comparison with Nvidia GPU-based solutions for the area.
 The main comparison metrics are trillion operations per second (TOPS) and efficiency which is measured as TOPS divided by power consumption.
-It is an overall nice basic overview of what is happening in the area with cool examples of 360-degree image construction and object classification:
-
-    \cite{Delmer:Mobileye:2019}.
+It is an overall nice basic overview of what is happening in the area with cool examples of 360-degree image construction and object classification.
 \end{itemize}
@@
 \begin{itemize}
-    \item Optical computing for AI moves closer to commercialization.
+    \item \cite{Wheeler:Optical-AI:2019}:
+
+    Optical computing for AI moves closer to commercialization.
 Startups originating from MIT and Princeton demonstrate interesting optical computing for AI solutions.
@@
 
-    \cite{Wheeler:Optical-AI:2019}.
+    \item \cite{Gwennap:Centaur-AI:2019}:
 
-    \item Centaur has developed an x86 processor with deep-learning accelerator --- Ncore.
+    Centaur has developed an x86 processor with deep-learning accelerator --- Ncore.
 The accelerator has a typical design for the field of application: it's a VLIW architecture relying on traditional SIMD approach with smaller sized vector elements like INT8, BF16, etc.
@@
 The article reports that so fat Centaur's software can only handle inference but not training.
-
-    \cite{Gwennap:Centaur-AI:2019}.
 \end{itemize}

### 2023-02-03 paper-reviews/papers-2023.tex
Add VMM trands article

@@
 
+\section*{Virtual Machines}
+\begin{itemize*}
+    \item \cite{Rosenblum:VMM-trends:2005}:
+
+    The article includes a very concise VMM history and some basic techniques like \textit{shadow paging}.
+Strict focus of the article is on VMware's solutions of that time.
+There described one interesting paravirtualization idea I didn't know about before --- \textit{ballon process} running inside a Guest OS and communicating with the VMM.
+The process can allocated a bunch of memory thus forcing the Guest OS to use its knowledge about other processes to page unused memory out.
+VMM thus can also page out this memory and then reclaim the memory used by the \textit{ballon process} - nice trick.
+\emd{itemize*}
+
 \end{document}

### 2023-02-03 paper-reviews/papers-2023.tex
Add review of FrobOS and fix several tiny problems

@@
 
-    The paper describes an early design of KVM?ARM port using the first version of ARM virtualization hardware support.
+    The paper describes an early design of KVM/ARM port using the first version of ARM virtualization hardware support.
 Authors claim to introduce a novel "split-mode" virtualization approach as a way to run hypervisor simultaneously in different CPU privileged modes.
@@
 The processor architecture supports INT8 and FP16 formats that are the current trend for DNN acceleration.
-On top of INT8 and FP16 the processor can be closes integrated with various accelerators.
-
+On top of INT8 and FP16 the processor can be closely integrated with various accelerators.
 \end{itemize}
@@
 Authors of the paper tried to address this important issue by implementing an automatic test case reducers for C.
+
+    \item \cite{Easson:FrobOS:2012}:
+
+    The article describes the ideas behind VMware's minimalistic test operating system FrobOS.
+Tbe operating system is a great instrument for bare-metal hardware tests.
+The tooling around the OS also enables nice and fast testing for their hypervisor: simple ways to run port and run the tests between different CPU and hypervisor modes.
 \end{itemize}
@@
 \section*{Virtual Machines}
-\begin{itemize*}
+\begin{itemize}
     \item \cite{Rosenblum:VMM-trends:2005}:
@@
 VMM thus can also page out this memory and then reclaim the memory used by the \textit{ballon process} - nice trick.
-\emd{itemize*}
+\end{itemize}
 

### 2023-02-03 paper-reviews/papers-2022.tex
Move paper reviews to a separate folder
Reformat paper overview from 2022

--- 2023-01-01 _posts/papers-2022.tex
@@
-\documentclass[12pt, a4paper, oneside]{article}
-
-\usepackage{hyperref}
-\hypersetup{colorlinks=true, allcolors=blue, pdfauthor=Evgenii Iuliugin}
-
-\usepackage{indentfirst}
-\usepackage[top=2.0cm, bottom=2.0cm, left=2.0cm, right=2.0cm,
-  footskip=2.0cm]{geometry}
-
-\pagestyle{empty}
-
-\usepackage[bibencoding=inputenc, backend=biber, language=auto, style=verbose]{biblatex}
-\newcommand{\bibpath}{../bibliography}
-\addbibresource{\bibpath/Machine Learning/hardware.bib}
-\addbibresource{\bibpath/Machine Learning/optimizations.bib}
-\addbibresource{\bibpath/Processors/riscv.bib}
-\addbibresource{\bibpath/Simulators/cycle-accurate.bib}
-\addbibresource{\bibpath/Simulators/description-languages.bib}
-\addbibresource{\bibpath/Simulators/functional.bib}
-\addbibresource{\bibpath/Simulators/frameworks.bib}
-\addbibresource{\bibpath/Simulators/target-analysis.bib}
-\addbibresource{\bibpath/compilers.bib}
-\addbibresource{\bibpath/embedded-systems.bib}
-\addbibresource{\bibpath/fun.bib}
-\addbibresource{\bibpath/performance.bib}
-\addbibresource{\bibpath/software-philosophy.bib}
-\addbibresource{\bibpath/virtual-machines.bib}
+\input{papers.tex}
 
@@
 
+A total of \total{citenum} papers read in the year.
+
 \section*{Machine Learning}
 \begin{itemize}
-    \item A new reconfigurable hardware architecture designed specifically to facilitate machine learning requirements:
+    \item \cite{Prabhakar:Plasticine:2017}:
 
-    \cite{Prabhakar:Plasticine:2017}.
+    A new reconfigurable hardware architecture designed specifically to facilitate machine learning requirements.
 
-    \item An optimization to increase parallelism of machine learning tasks:
+    \item \cite{Yuanzhong:Cross-Replica-Sharding:2020}:
 
-    \cite{Yuanzhong:Cross-Replica-Sharding:2020}.
+    An optimization to increase parallelism of machine learning tasks.
 \end{itemize}
@@
 \begin{itemize}
-    \item Western Digital builds its own RISC-V processor:
+    \item \cite{Wheeler:WD-RISC-V:2019}:
 
-    \cite{Wheeler:WD-RISC-V:2019}.
+    Western Digital builds its own RISC-V processor.
 \end{itemize}
@@
 \begin{itemize}
-    \item Basic hardware performance modeling approach with separation of functional and timing simulation using ports:
+    \item \cite{Emer:Asim:2002}:
 
-    \cite{Emer:Asim:2002}.
+    Basic hardware performance modeling approach with separation of functional and timing simulation using ports.
 
-    \item Micro architectural processor simulation framework:
+    \item \cite{Hughes:Rsim:2002}:
 
-    \cite{Hughes:Rsim:2002}.
+    Micro architectural processor simulation framework.
 
-    \item Infrastructure and framework for cycle-accurate and functional simulation:
+    \item \cite{Austin:SimpleScalar:2002}:
 
-    \cite{Austin:SimpleScalar:2002}.
+    Infrastructure and framework for cycle-accurate and functional simulation.
 
-    \item KVM-based direct execution for ARM:
+    \item \cite{Junger:ARM-on-ARM:2020}:
 
-    \cite{Junger:ARM-on-ARM:2020}.
+    KVM-based direct execution for ARM.
 
-    \item Overview of Simics simulator:
+    \item \cite{Magnusson:Simics:2002}:
 
-    \cite{Magnusson:Simics:2002}.
+    Overview of Simics simulator.
 
-    \item Basic algorithm for multiprocessor target on multiprocessor host simulation:
+    \item \cite{Kenney:MP-on-MP:2015}:
 
-    \cite{Kenney:MP-on-MP:2015}.
+    Basic algorithm for multiprocessor target on multiprocessor host simulation.
 
-    \item Single instruction JIT compilation approach:
+    \item \cite{Nohl:JIT-Simulation:2002}:
 
-    \cite{Nohl:JIT-Simulation:2002}.
+    Single instruction JIT compilation approach.
 
-    \item Memory and thread usage analysis for software running inside a virtual platform:
+    \cite{Dahl:VP-Sanitizer:2018}:
 
-    \cite{Dahl:VP-Sanitizer:2018}.
+    \item Memory and thread usage analysis for software running inside a virtual platform.
 \end{itemize}
@@
 \begin{itemize}
-    \item An amazing overview of basic but demonstrative optimizations performed by a modern C/C++ compiler:
+    \item \cite{Godbolt:Optimizations:2020}:
 
-    \cite{Godbolt:Optimizations:2020}.
+    An amazing overview of basic but demonstrative optimizations performed by a modern C/C++ compiler.    
 \end{itemize}
@@
 \begin{itemize}
-    \item PhD thesis by Jakob Engblom focused on studying of worst case execution time analysis for embedded systems:
+    \item \cite{Engblom:WCET:2002}:
 
-    \cite{Engblom:WCET:2002}.
+    PhD thesis by Jakob Engblom focused on studying of worst case execution time analysis for embedded systems.
 \end{itemize}
@@
 \begin{itemize}
-    \item Really fun connection between texts of several popular songs and computer basics:
+    \item \cite{Colwell:Rock-N-Roll:2002}:
 
-    \cite{Colwell:Rock-N-Roll:2002}.
+    Really fun connection between texts of several popular songs and computer basics.
 \end{itemize}
@@
 \begin{itemize}
-    \item A way to determine maximum performance of a particular system together with few ideas on how to decide what optimizations make sense to do and when:
+    \item \cite{Williams:Roofline:2009}:
 
-    \cite{Williams:Roofline:2009}.
+    A way to determine maximum performance of a particular system together with few ideas on how to decide what optimizations make sense to do and when.
 \end{itemize}
@@
 \begin{itemize}
-    \item Discussions on what a perfect design process show look like:
+    \item \cite{Parnas:Fake-It:1986}:
 
-    \cite{Parnas:Fake-It:1986}.
+    Discussions on what a perfect design process show look like.
 
-    \item Philosophical discussions about causes of errors in software:
+    \item \cite{Horner:Error-Rates:2019}:
 
-    \cite{Horner:Error-Rates:2019}.
+    Philosophical discussions about causes of errors in software.
 \end{itemize}
@@
 \begin{itemize}
-    \item A set of techniques to detect a virtualized Android environment using basic phone identification information, sensor data and timing of processor instruction execution:
+    \item \cite{Petsas:Android-VM-Detection:2014}:
 
-    \cite{Petsas:Android-VM-Detection:2014}.
+    A set of techniques to detect a virtualized Android environment using basic phone identification information, sensor data and timing of processor instruction execution.
 \end{itemize}

### 2023-02-07 paper-reviews/papers-2023.tex
Add SystemC and gem5 integration paper

@@
 Some of the links in references are also worse looking through later.
+
+    \item \cite{Menard:SystemC-gem5:2017}:
+
+    The article provides a short summary of transaction-based communication in SystemC TLM and in gem5.
+However the main focus of the paper is on an adapter layer between SystemC TLM and gem5 transactions.
+The layer provides bi-directional translation for the transaction.
+The layer was tested on a special-purpose worse-case setup and a couple of more realistic workflows.
+On the worst-case scenario the layer demonstrated roughly 2x slowdown compared to SystemC only.
+The integration is claimed to be merged to gem5 official repo.
 \end{itemize}

### 2023-02-10 paper-reviews/papers-2023.tex
Add Full Speed Ahead paper

@@
 The integration is claimed to be merged to gem5 official repo.
+
+    \item \cite{Sandberg:FullSpeedAhead:2015}:
+
+    The article describes an approach for fast full-system cycle-accurate simulation to enable efficient hardware and software co-design.
+The approach address fixed-software limitation of the standard checkpoint-based critical path cycle-accurate simulation approach.
+First, the authors reduced the porting of functional warming for the samples and introduced error estimation algorithm for cache warming.
+The error estimation compares best- and worst-case results (run in parallel) to determine whether limited functional warming has significant effect on the results or not.
+The technique is named Full Speed Ahead.
+The authors extended the technique with parallel sample execution.
+The main idea of the parallel Full Speed Ahead approach is to break target workload into samples and execute the samples in parallel while fast-forwarding to the sample boundaries using hardware-supported virtualization.
+As result, the authors claims to achieve near-native simulation speed while running on a wide enough (in terms of processor cores) host systems.
 \end{itemize}

### 2023-02-12 paper-reviews/papers-2023.tex
Add DSP Derby article

@@
 On top of INT8 and FP16 the processor can be closely integrated with various accelerators.
+
+    \item \cite{Eyre:DSP-Derby:2001}:
+
+    The article presents digital signal processor (DSP) design and development challenges.
+The author compares techniques used in DSPs with techniques typical for general-purpose processors.
+The article emphasizes DSP's requirement for predictable executing speed that is often more important than performance.
+The requirement is of course harder to fulfill with caches and out-of-order execution.
+VLIW, SIMD and highly compressed encodings are mentioned as an example of techniques prevailing in DPSs.
+There is also performance comparison of some DSPs with a superscalar processor that is based on FFT and FIR filter benchmarks.
 \end{itemize}

### 2023-02-16 paper-reviews/papers-2023.tex
Add BT future article

@@
 
+\section*{Binary Translation}
+\begin{itemize}
+    \item \cite{Altman:BT-Future:2001}:
+
+    The article provides a nice overview of Transmeta Crusoe and IBM DAISY with some insides on the special-purpose underlying hardware.
+There is an overview of other binary translation systems including LaTTe JVM.
+There is a clear focus on speculative operations to be performed by Crusoe and DAISY binary translators.
+One optimization had a bit unclear benefits when I read the article --- the speculative loads performed by DAISY.
+While reading the article I only thought about the speculative operation as a cache warmup for the later reload.
+The problem here is that speculative loads could overlap with stores performed before the actual load should occur.
+To address this issue DAISY uses a \textbf{load-verify} operations that is insert to the location of the original load.
+The operation reloads the memory value and compares it with the speculatively pre-loaded one.
+Depending on the result DAISY either continues as normal or enters recovery mode.
+While writing the note I realized that the load speculation might allow other speculations to be performed and thus the recovery mode is needed to roll them back if the load is overlapped.
+The article points to the \href{https://patents.google.com/patent/US5758051A}{US5758051A} patent with more details that could be worse to read later.
+\end{itemize}
+
 \end{document}

### 2023-02-24 paper-reviews/papers-2023.tex
Add reference to CPU top down analysis

@@
 
+\section*{Performance}
+\begin{itemize}
+    \item \cite{Yasin:TopDownCPUPerfAnalysis:2014}:
+
+    The paper proposes a top down performance analysis method for superscalar out-of-order processors.
+The paper discusses a couple of challenges for low-level CPU performance analysis; problems and limitation of the existing methods.
+The paper suggest a tree-like structure for the analysis.
+The top level of the hierarchy includes front end, bad speculations, retiring and frontend.
+Looking at specific part of the hierarchy allows to disregard events happening at other parts thus narrowing down the problem while moving to the leaves of the hierarchy.
+The author presented few usage examples of the methods and proposed changes to Intel's performance monitoring unit architecture.
+\end{itemize}
+
 \end{document}

### 2023-02-27 paper-reviews/papers-2023.tex
Add TensorFlow paper

@@
 The article reports that so fat Centaur's software can only handle inference but not training.
+
+    \item \cite{Abadi:TensorFlow:2016}:
+
+    The article describes ideas and reasons that stay behind TensorFlow.
+Nowadays one can hardly imaging machine learning without TensorFlow.
+All the accelerators I know about have support for TensorFlow or aim to support it.
+To represent computations, TensorFlow uses a single dataflow graph where vertices represent operations (atomic units of computations) and edges represent input and output to a vertex.
+The dataflow graph can be broken into subgraphs and distributed between multiple servers for execution.
+Basic control flow is supported using \texttt{Switch} and \texttt{Merge} operations.
+The system provides support for CPUs, GPUs, Google TPUs and allows extensions for other hardware systems.
 \end{itemize}

### 2023-03-12 paper-reviews/papers-2023.tex
Add new paper description

@@
 Industry practice so far showed that VLIW could be an amazing architecture for some DSPs while a superscalar processor delivers better results for general purpose usage.
+
+    \item \cite{Rajwar:SpeculativeLockElision:2001}:
+
 \end{itemize}
@@
 The article points to the \href{https://patents.google.com/patent/US5758051A}{US5758051A} patent with more details that could be worse to read later.
+
+    \item \cite{Hamayun:StaticBT-VLIW:2013}:
+
+    One of the most confusing papers I read.
+Unclear why KVM is described in the paper what is the so <<native>> about that approach.
+I can see two well-known ideas in the paper --- cross-platform compilation and static binary translation.
+Main focus of the paper is on static binary translation for VLIW CPUs.
+There is a good description of VLIW weirdness like register dependencies withing the package and delay slots.
+The authors built a static binary translation that is capable of handling all the tricky VLIW peculiarities.
 \end{itemize}

### 2023-03-14 paper-reviews/papers-2023.tex
Add description for SLE paper

@@
 
+    The paper describes Speculative Lock Elision --- CPU microarchitectural technique aimed to support transaction memory.
+The proposed technique suggests automatic detection and elision of lock acquire and release instructions if and only if a processor can guarantee atomic memory updates between the locks.
+No ISA level changes are required.
+The existing cache coherency mechanism can to be used to guarantee atomicity.
+The technique's behavior is similar to how a branch predictor speculation works --- assume that lock can be elided, elide the lock, roll-back if the assumption turned out to be wrong, commit the results otherwise.
+The roll-back can be achieved either using reorder buffer (ROB) or register checkpointing.
+The later allows larger critical sections to be speculated.
+Speculative Lock Elision supports nested locks --- only one (can be the outer one) of the locks is elided in that case.
+Intel has implemented the technique in its Hardware Lock Elision (HLE) extension that is well known for its bugginess.
+HLE is now deprecated and removed from the latest processors.
 \end{itemize}

### 2023-03-15 paper-reviews/papers-2023.tex
Add parallel JIT paper

@@
 The authors built a static binary translation that is capable of handling all the tricky VLIW peculiarities.
+
+    \item \cite{Qin:ParallelJIT:2006}:
+
+    The paper proposes a parallel JIT-compilation technique for ISSes. The idea behind technique is rather primitive. Here is a short summary:
+
+    \begin{enumerate}
+        \item Interpreter runs and collects statistics about block usage.
+        \item Once a hot block is detected it is added to compilation queue.
+        \item Compilation workers run in parallel host threads. C++ code is generated for each block. GCC is used to compile the blocks to DLLs.
+        \item DLLs are loaded and used to simulate the block.
+    \end{enumerate}
+
+    Self-modifying code is supported through monitoring of writes to compiled blocks.
+The authors also added a DLL cache to avoid compilation overhead for subsequent runs of the same simulation.
+Easy to read article that would be a good reference for \herf{https://github.com/yulyugin/sim-lectures}{<<Fundamentals of Full-Platform Simulation>>} MIPT course.
 \end{itemize}

### 2023-03-15 paper-reviews/papers-2023.tex
Add paper about static analysis at facebook

@@
 
+\section*{Static Analysis}
+\begin{itemize}
+    \item \cite{Distefano:StaticAnalysisFB:2019}:
+
+    The paper presents Infer and Zoncolan (security analysis) --- static analysis tools developed and used by Facebook.
+On top of standard false positive and false negative metrics authors added fix rate metric measuring the proportion of reported issues that were actually fixed.
+One of the observation is that the fix rate gets closer to 70\% when the static analysis is applied during diff (pull request) time.
+Meanwhile, the fix rate is near zero when static analysis is applied once on the entire code base.
+The overall approach to development of the static analysis tools looks very reasonable to me including the initial ideas, integration and addition of new rules.
+I would definitely recommend to read this paper for anyone who is going to work on custom static analysis tools.
+So far the approach looks to be the most reasonable static analysis application I saw in the industrial projects I worked on.
+Infer is available as open source and I should definitely try it on current projects.
+\end{itemize}
+
 \end{document}

### 2023-03-20 paper-reviews/papers-2023.tex
Add scalar waving paper

@@
 
+\section*{GPU}
+\begin{itemize}
+    \item \cite{Yilmazer:ScalarWaving:2014}:
+
+    The authors discovered that GPU programs contain significant portion of scalar instructions.
+30\% for the GPU selected benchmarks in the paper.
+The papers states that scalar instructions are typically executed as SIMD with all vector elements having the same values.
+Thus, the same computation is performed redundantly.
+One of the solution for the problem is to run the calculations once and than broadcast the results in order to reduce power consumption on unnecessary operations.
+The authors suggested to statically detect scalar operations during compilations.
+One negative effect of this is reduced utilization of SIMD lanes.
+To improve utilization the  paper proposed scalar waving --- grouping of same scalar instructions processing different scalars into a single SIMD instruction.
+This grouping is suggest to be performed during runtime through by GPU itself.
+Further, the authors developed simultaneous scalar and SIMD group waving --- grouping and a SIMD instructions and scalar ware.
+GPU needs to be able to execute two different operations as a part of SIMD instruction in this case.
+The authors claim to achieve up to 25\% percent performance improvement for the selected benchmarks.
+The performance studies are performed on \href{https://github.com/gpgpu-sim}{GPGPU-Sim} simulator.
+\end{itemize}
+
 \end{document}

### 2023-03-20 paper-reviews/papers-2023.tex
Add review of Turbo Boost evaluation paper

@@
 
-    The article presents digital signal processor (DSP) design and development challenges.
+    The article presents digital signal processor's (DSP) design and development challenges.
 The author compares techniques used in DSPs with techniques typical for general-purpose processors.
@@
 There is also performance comparison of some DSPs with a superscalar processor that is based on FFT and FIR filter benchmarks.
+
+    \item \cite{Charles:TruboBoostEvaluation:2009}:
+
+    The article presents evaluation of Intel's Turbo Boost technology first introduce in Nehalem microarchitecture.
+The technology may increase the processor's frequency when load and temperature are low.
+SPEC CPU2006 is used as benchmark for the evaluation.
+Benchmark suites are classified as CPU-intensive and memory-intensive based on LLC cache-miss rate.
+The authors perform three kind of tests:
+
+    \begin{enumerate}
+        \item Isolation tests --- every benchmark runs in isolation with and without Turbo Boost to measure speedup that can be achieved through the frequency increase.
+        \item Paired tests --- pairs of the benchmarks run simultaneously to test if the process can still make effective use of Turbo Boost.
+        \item Saturation tests --- all processor cores are loaded with different combination of CPU- and memory-intensive benchmarks to test whether Turbo Boost activates when all cores are loaded.
+    \end{enumerate}
+
+    One interesting observation from the results is that for saturation tests maximum Turbo Boost activation is not always corresponds to the best overall performance of the system.
+
+    Another set of tests described in the paper is multi-threaded workloads from PARSEC and BLAST benchmark suits.
+The idea behind the tests is also pretty interesting: according to Amdahl's law speedup of parallel application is limited by its sequential component.
+The authors check whether Amdahl's law can be mitigated through the frequency increase during the sequential execution.
+There is a very interesting citation about that in the paper that is worse looking at later and is probably a good source for \href{https://github.com/yulyugin/mipt-parallel-computing}{<<MIPT parallel computing>> course}.
 \end{itemize}

### 2023-03-30 paper-reviews/papers-2023.tex
Add software fault isolation paper

@@
 
+\section*{Fault Isolation}
+\begin{itemize}
+    \item \cite{Wahbe:SoftwareFaultIsolation:1993}:
+
+    The authors proposed a software technique aimed to isolate faults in third party software modules loaded as extension.
+The idea is to run both the main module and the submodules in one address space to avoid expansive RPC communication.
+Fault isolation is supposed to be achieved through binary patching which is implemented on compile time by the authors.
+Each modules has a dedicated part of the shared memory space.
+The modules are not supposed to access other module's memory regions.
+To achieve that every memory access or jump that uses indirect address is wrapped by a special instruction sequence ensuring isolation.
+Few registers are considered to be special purpose registers and thus reserved for use by the isolation mechanism.
+Trusted modules like the core module do not need to be modified and thus avoid the patching overhead.
+Intra module communication is supposed to be done through an intermediate memory space using special call mechanism.
+Operation is system should be modified to protect one module from closing resources like file descriptors of another module.
+The authors report one to two orders of magnitude performance improvement compared to existing at that moment RPC mechanisms.
+\end{itemize}
+
 \end{document}

### 2023-03-31 paper-reviews/papers-2023.tex
Add HTT validation paper

@@
 The tooling around the OS also enables nice and fast testing for their hypervisor: simple ways to run port and run the tests between different CPU and hypervisor modes.
+
+    \cite{IntelHTTValidation}:
+
+    The article shares a story and decisions made during Intel's Hyper Threading Technology pre-silicon validation.
+Some implementation details of the technology are described as well.
+The amount of tests that needs to be performed explodes when going from single logical core system to the hyper threaded system.
+All the state now falls into four categories: duplicated, partitioned, entry tagged and fully shared.
+The four state categories together with new modes of operation lead to test explosion.
+One of the interesting solutions was to switch between logical threads for regular testing instead of of always testing both.
+The authors also made an interesting post mortem analysis of both issues discovered during pre-silicon validation and the issues found in silicon.
 \end{itemize}

### 2023-04-02 paper-reviews/papers-2023.tex
Fix typos

@@
 
-    \cite{IntelHTTValidation}:
+    \cite{Burns:IntelHTTValidation:2002}:
 
@@
 The authors also added a DLL cache to avoid compilation overhead for subsequent runs of the same simulation.
-Easy to read article that would be a good reference for \herf{https://github.com/yulyugin/sim-lectures}{<<Fundamentals of Full-Platform Simulation>>} MIPT course.
+Easy to read article that would be a good reference for \href{https://github.com/yulyugin/sim-lectures}{<<Fundamentals of Full-Platform Simulation>>} MIPT course.
 \end{itemize}

### 2023-04-04 paper-reviews/papers-2023.tex
Add dynamic warp formation paper

@@
 
-    \cite{Burns:IntelHTTValidation:2002}:
+    \item \cite{Burns:IntelHTTValidation:2002}:
 
@@
 The performance studies are performed on \href{https://github.com/gpgpu-sim}{GPGPU-Sim} simulator.
+
+    \item \cite{Fung:DynamicWrapGPUControlFlow:2007}:
+
+    The paper is focused on GPGPU control flow problems that represent general control flow issues with SIMD --- what should be done if control flow diverges for SIMD lines.
+The paper describes several possible solutions for the problem: scalar execution (more typical for CPUs), reconvergence at the immediate post-dominator, dynamic warp formation that is supposed to improve SIMD pipeline utilization.
+The authors show that reconvergence at the immediate post-dominator is not always the best choice.
+They propose an implementation of dynamic warp formation together with several heuristics that are supposed to define when the formed warps should be pushed for execution.
+The authors claim that dynamic warp formation improves performance by 20.7\% on average compared to reconvergence at the immediate post-dominator.
+They estimated that area increase because of the proposed hardware changes to be around 4.7\%.
 \end{itemize}

### 2023-04-06 paper-reviews/papers-2023.tex
Add MARSS reference

@@
 As result, the authors claims to achieve near-native simulation speed while running on a wide enough (in terms of processor cores) host systems.
+
+    \item \cite{Patel:MARSS:2011}:
+
+    The article is about Micro Architectural and System Simulator (MARSS).
+The simulator is constructed as a combination of QEMU and PTLSim.
+PTLSim and QEMU share the same processor structure which allows them to switch seamlessly thus providing a gear-shit capability for mixing functional and cycle-accurate execution.
+QEMU provides all the peripherals in the setup.
+The article uses term emulation for function models and simulation for cycle-accurate models.
+This is another view on this controversial topic that can be a reference for the \href{https://github.com/yulyugin/sim-lectures}{<<Fundamentals of Full-Platform Simulation>>} course.
 \end{itemize}

### 2023-04-24 paper-reviews/papers-2023.tex
Add MPI communication paper

@@
 
+\section*{Parallel Programming}
+\begin{itemize}
+    \item \cite{Thakur:MPIOptimization:2005}:
+
+    The article presents different high-performance implementations of collective MPI operations.
+The authors assume that the time to send similar package between any two nodes is the same and does not depend on number of concurrent communications.
+This is a very simple cost model but seems to be sufficient for the author's needs.
+The article is a good source for different one-to-all, all-to-one, all-to-all communication algorithms.
+Interesting but expected idea from the article is to use different communication algorithms depending on message size.
+\end{itemize}
+
 \end{document}

### 2023-06-09 paper-reviews/papers-2023.tex
Add new papers

@@
 This is another view on this controversial topic that can be a reference for the \href{https://github.com/yulyugin/sim-lectures}{<<Fundamentals of Full-Platform Simulation>>} course.
+
+    \item \cite{Moore:Imperas-Custom-RISCV:2019}:
+
+    The article describes a standard way of adding new processor instructions purely based on instruction count: find a pattern, add instruction, makes sure it works, measure instruction count to estimate how performance changes, repeat it needed.
+The approach is applied to RISC-V processors using IMPERAS ISS.
+Could be a good reference for students who are not aware how this could be done.
 \end{itemize}
@@
 
-\section*{Fault Isolation}
+\section*{Software Attacks}
 \begin{itemize}
@@
 The authors report one to two orders of magnitude performance improvement compared to existing at that moment RPC mechanisms.
+
+    \item \cite{Murdock:Plundervolt:2020}:
+
+    The authors discovered an attack vector to Intel SGX through processor undervolting which is available on modern processors via undocumented MSRs.
+In Intel SGX's model, both OS and physical access are considered to be untrusted.
+In my opinion this is enough to make the technology rather questionable.
+This paper proves my doubts about the technology.
+Intel's response to the problem was quite expected - simply hide the undocumented voltage control MSRs thus hiding the problem without actually fixing it.
+I admit, it must be a pain to protect against such kind of attacks.
+Microsoft has a nice software guarding against overflows that may be in this particular case caused by faulty multiplication --- divide the result and compare it with then incoming operands.
 \end{itemize}

### 2023-07-19 paper-reviews/papers-2023.tex
Add CGC Monitor reference

@@
 Could be a good reference for students who are not aware how this could be done.
+
+    \item \cite{Thompson:CGCMonitor:2018}:
+
+    The article is a good demonstration of what can be achieved using analysis capabilities provided by a simulation system.
+Authors of the paper built vetting system for a cyber challenge.
+The main purpose of the system was to make sure that none of the competitors is trying to subvert the competition infrastructure.
+The vetting system was build on top of Simics hap and breakpoint capabilities thus providing non-intrusive analysis with support for reverse execution.
 \end{itemize}

### 2023-10-04 paper-reviews/papers-2023.tex
Add FP8 reference

@@
 The system provides support for CPUs, GPUs, Google TPUs and allows extensions for other hardware systems.
+
+    \item \cite{Micikevicius:FP8ForDL:2022}
+
+    The article describes two new 8-bit floating point formats: E4M3 and E5M2.
+The formats are a natural progression of the current trend of data type shrinking that is aimed to improve computational efficiency of machine learning training and inference.
+One of the formats --- E5M2 --- can be trivially converted to FP16.
+In a similar way that bfloat16 can be converted to FP32.
+The authors show that results of training and inference done using the new formats effectively matches the results achieved by 16-bit formats.
 \end{itemize}

### 2023-10-30 paper-reviews/papers-2023.tex
Add attention paper
Add FlashAttention reference

@@
 
-    \item \cite{Micikevicius:FP8ForDL:2022}
+    \item \cite{Micikevicius:FP8ForDL:2022}:
 
@@
 The authors show that results of training and inference done using the new formats effectively matches the results achieved by 16-bit formats.
+
+    \item \cite{Vaswani:Attention:2023}:
+
+    The article proposes the Transformer machine learning network architecture that is solely based on attention mechanisms.
+The authors show that the new models are superior compared to recurrent or conventional neural networks when applied to machine translation tasks.
+
+    \item \cite{Dao:FlashAttention:2022}:
+
+    The article proposes FlashAttention, an IO-aware exact attention algorithm focused on reducing of read/write accesses between GPU on-chip SRAM and high bandwidth memory (HBM).
+The authors analyze IO complexity of the algorithm, showing that it required fewer HBM accesses than standard attention algorithm.
+The article also extends FlashAttention to block-sparse attention, an approximate attention algorithm.
+FlashAttention algorithm is based on kernel fusion (most common approach to accelerate memory-bound operations by loading data from HDM once and applying it to several operators, instead of loading it multiple times for each operator) and tiling (computation using blocks instead of the full dataset).
+The authors demonstrated both training speed and accuracy/perplexity improvements because of larger context allowed by faster training.
 \end{itemize}

### 2023-11-13 paper-reviews/papers-2023.tex
Add ASTra-Sim reference

@@
 
-\section*{Parallel Programming}
+\section*{Collective Communication}
 \begin{itemize}
@@
 Interesting but expected idea from the article is to use different communication algorithms depending on message size.
+
+    \item \cite{Rashidi:AstraSim:2020}:
+
+    The article present ASTra-Sim --- simulator and modeling approach aimed to navigate SW/HW design space of distributed deep learning training platforms.
+The paper justifies the need of such simulation defining some basic concepts used in the field.
+ASTra-Sim is based on Garnet (part of GEM5) network simulator and uses event-driven execution model.
+ASTra-Sim implements topology-aware collective operations and different parallelism approaches of training.
+ASTra-Sim provides interfaces for compute and network simulators allowing to replace the default implementation with other simulators.
+The research is focused on network topologies and have description for how different communication operations can be mapped on the studied topologies.
+The study is focused on torus and all-to-all topologies and hierarchical versions of the above.
 \end{itemize}

### 2023-11-21 paper-reviews/papers-2023.tex
Add note about compute
Add ASTRA-sim2.0 paper

@@
 The authors demonstrated both training speed and accuracy/perplexity improvements because of larger context allowed by faster training.
+
+    \item \cite{Rashidi:AstraSim:2020}:
+
+    The article present ASTra-Sim --- simulator and modeling approach aimed to navigate SW/HW design space of distributed deep learning training platforms.
+The paper justifies the need of such simulation defining some basic concepts used in the field.
+ASTra-Sim is based on Garnet (part of GEM5) network simulator and uses event-driven execution model.
+ASTra-Sim implements topology-aware collective operations and different parallelism approaches of training.
+ASTra-Sim provides interfaces for compute and network simulators allowing to replace the default implementation with other simulators.
+The research is focused on network topologies and have description for how different communication operations can be mapped on the studied topologies.
+The study is focused on torus and all-to-all topologies and hierarchical versions of the above.
+For compute part the authors consider TPU-like hardware and only take GEMM modeling into account (all other operators considered to take parametrized delay).
+
+    \item \cite{Won:ASTRA-sim2.0:2023}:
+
+    The article presents extension to ASTRA-sim simulator that is aimed to navigate SH/HW design space of distributed deep learning training platforms.
+Compared to the original design of ASTRA-sim the authors switch from basically being cycle-accurate network simulator to an analytical performance simulator.
+The simulator splits distributed training into three layers: workload, system and network.
+Multi-dimensional network designs based on three basic topologies (ring, fully connected and switched) are supported.
+Compute simulation is done in roofline matter.
+Memory disaggregation modelling (fundamental solution to overcome capacity of a single unit) is added to the simulator.
 \end{itemize}
@@
 Interesting but expected idea from the article is to use different communication algorithms depending on message size.
-
-    \item \cite{Rashidi:AstraSim:2020}:
-
-    The article present ASTra-Sim --- simulator and modeling approach aimed to navigate SW/HW design space of distributed deep learning training platforms.
-The paper justifies the need of such simulation defining some basic concepts used in the field.
-ASTra-Sim is based on Garnet (part of GEM5) network simulator and uses event-driven execution model.
-ASTra-Sim implements topology-aware collective operations and different parallelism approaches of training.
-ASTra-Sim provides interfaces for compute and network simulators allowing to replace the default implementation with other simulators.
-The research is focused on network topologies and have description for how different communication operations can be mapped on the studied topologies.
-The study is focused on torus and all-to-all topologies and hierarchical versions of the above.
 \end{itemize}

### 2024-02-19 paper-reviews/papers-2024.tex
Add Hyena paper

\input{papers.tex}

\begin{document}

A total of \total{citenum} papers read in the year.

\section*{Machine Learning}
\begin{itemize}
    \item \cite{Poli:Hyena:2023}:

    The paper proposes Hyena --- a subquadratic replacement for attention constructed using long convolutions (FFT in particular). Hyena is aiming to solve Transformer scaling problem. Transformers are currently domination in deep learning. However, Transformers are based on the attention operator that exhibits quadratic computational complexity in sequence length which in turn is limiting context available to the model. This fact makes breaking the quadratic barrier a key step forwards new deep learning capabilities.
\end{itemize}

\end{document}

### 2024-05-16 paper-reviews/papers-2024.tex
Add FlashFFTConv paper

@@
 This fact makes breaking the quadratic barrier a key step forwards new deep learning capabilities.
+
+    \item \cite{Fu:FlashFFTConv:2024}:
+
+    FlashFFTConv algorithm proposed in the paper tries to address major bottlenecks related to convolutional models with long filters --- poor utilization of specialized matrix multiplication units and expensive I/O between layers of memory hierarchy.
+The major bottleneck is caused by the Fast Fourier Transform.
+Convolutional models lag behind the most Transformers even though FFT allows to run in O(N log N) instead of  O(N^2) common Transformers.
+FlashFFTConv uses a matrix decomposition that computes FFT using matrix multiplication units and enables kernel fusion for long sequences that in turn allows to reduce the expensive I/O operations.
 \end{itemize}

### 2024-05-31 paper-reviews/papers-2024.tex
Add Monarch paper

@@
 FlashFFTConv uses a matrix decomposition that computes FFT using matrix multiplication units and enables kernel fusion for long sequences that in turn allows to reduce the expensive I/O operations.
+
+    \item \cite{Dao:Monarch:2022}:
+
+    The paper proposes a class of hardware efficient Monarch matrices.
+The matrices are proposed as a replacement for dense weight matrices in order to reduce compute/memory requirements of large neural networks.
+Monarch matrices are block diagonal matrices that can represent many fast transforms such as Fourier or convolution.
+The authors show that dense-to-sparse conversion has an analytically optimal solution.
+A Monarch matrix is described by $2n\sqrt{n}$ parameters instead $n^2$ for dense matrix.
+Monarch parametrization allows efficient utilization of batch matrix multiply units available in contemporary hardware such as GPUs.
+The parametrization allows to achieve up to $2x$ speedup compared to dense matrix multiply.
+Algorithm proposed in the paper allows to fine-tune a pre-trained model into a model with Monarch weight matrices.
+The work presented in the paper is a step towards doing structural matrices more practical.
 \end{itemize}

### 2024-05-31 paper-reviews/papers-2024.tex
Add Megatron-LM paper

@@
 The work presented in the paper is a step towards doing structural matrices more practical.
+
+    \item \cite{Shoeybi:MegatronLM:2020}:
+
+    The article describes techniques for training large transformer models that is based on intra-layer model parallel approach.
+The approach doesn't require a complier or library changes and is only based on PyTorch level changes to a model.
+The approach is demonstrated by increasing number of parameters for existing transformer models.
+The main idea of the technique is to partition GEMM in a way that allows non-linear functions to be applied to each partition independently.
+The technique is aimed to reduce number of communications (reductions) introduces py partitioning.
+The technique is orthogonal to data parallelism thus allowing combination of the two.
+The approach has demonstrated near linear scaling.
 \end{itemize}

### 2024-06-14 paper-reviews/papers-2024.tex
Add microscaling paper

@@
 The approach has demonstrated near linear scaling.
+
+    \item \cite{Rouhani:Microscaling:2023}:
+
+    The paper investigates performance on narrow-bit data formats for machine learning applications.
+Narrow-bit data types are aimed to reduce computational and storage costs of the applications.
+Mixed precision calculations are considered in the paper: GeMMs performed in the narrow-bit data format and all the vector calculations are preformed in Bfloat16 or FP32.
+The authors showed that both model training and model inference performance doesn't downgrade that much when going to narrow-bit formats.
+INT8 almost matched the baseline FP32 in experiments done by the paper.
+The results demonstrated that training with MXFP4 weights and MXFP6 activations incurs only a minor penalty in model performance loss.
 \end{itemize}

### 2024-07-09 paper-reviews/papers-2024.tex
Add BERT paper

@@
 The results demonstrated that training with MXFP4 weights and MXFP6 activations incurs only a minor penalty in model performance loss.
+
+    \item \cite{Devlin:Bert:2019}:
+
+    The paper presents BERT --- language model based on multi-layer bidirectional transformer.
+BERT training consists of two parts: pre-training and fine-tuning.
+Pre-training is done on unlabeled data.
+For fine-tuning the model is first initialized with pre-trained parameters and then all the parameters are fine-tuned using labeled data from the target tasks.
+Fine-tuning is relatively inexpensive compared to pre-training.
+BERT pre-training is done using "masked LM" approach --- random words in a text sequenced are masked and the language model is trained to predict them correctly thus allowing to obtain bidirectional model.
+To understand dependencies between sentences, the model is trained on labeled sentence pairs that are related (second sentence is following the first one in the original text) and labeled as "next" in 50\% cases and not related (second sentence is random) and labeled as "not next" in the other 50\% cases.
 \end{itemize}

### 2024-07-19 paper-reviews/papers-2024.tex
Add FlashAttention-V2

@@
 To understand dependencies between sentences, the model is trained on labeled sentence pairs that are related (second sentence is following the first one in the original text) and labeled as "next" in 50\% cases and not related (second sentence is random) and labeled as "not next" in the other 50\% cases.
+
+    \item \cite{Dao:FlashAttentionV2:2023}:
+
+    The paper presents an improved version of Flash Attention algorithm.
+The main change include: 1.
+reduced number of non-matmul FLOPs (matmul throughput can be up to 16X higher than non-matmul throughput); 2.
+parallelization along sequence length dimension in both forward and backward passes; 3.
+work partitioning aimed to reduce communication and shared memory accesses.
+FlashAttentionV2 is claimed to be around 2X faster than FlashAttention.
 \end{itemize}

### 2024-08-08 paper-reviews/papers-2024.tex
Fix latex build
Add dropout paper

@@
 The major bottleneck is caused by the Fast Fourier Transform.
-Convolutional models lag behind the most Transformers even though FFT allows to run in O(N log N) instead of  O(N^2) common Transformers.
+Convolutional models lag behind the most Transformers even though FFT allows to run in $O(N log N)$ instead of  $O(N^2)$ common Transformers.
 FlashFFTConv uses a matrix decomposition that computes FFT using matrix multiplication units and enables kernel fusion for long sequences that in turn allows to reduce the expensive I/O operations.
@@
 FlashAttentionV2 is claimed to be around 2X faster than FlashAttention.
+
+    \item \cite{Hinton:Dropout:2012}:
+
+    The paper described dropout training approach aimed to prevent co-adaptation between several feature detectors within a network.
+The co-adaptation often leads to overfitting of the network.
+This problem is greatly reduced by randomly omitting results of some (half in the paper) feature detectors.
+The dropout ensures that each neuron learns to detect a feature in isolation that is in turn improves network's performance as demonstrated by the paper.
 \end{itemize}

### 2024-08-27 paper-reviews/papers-2024.tex
Add a paper about linguistic regularities in word representation

@@
 The dropout ensures that each neuron learns to detect a feature in isolation that is in turn improves network's performance as demonstrated by the paper.
+
+    \item \cite{Mikolov:LinguisticRI:2013}:
+
+    The paper examines how linguistic regularities are present in vector-space word representations implicitly learned by language models.
+The words are represented as high dimensional real valued vectors.
+The study shows that syntactic and semantic language regularities (male/female, singular/plural, etc.) are characterized by a relation-specific vector offset.
+The authors showed that it is possible to use the regularities to answers analogy questions.
 \end{itemize}

### 2024-09-30 paper-reviews/papers-2024.tex
Add MQA paper

@@
 The authors showed that it is possible to use the regularities to answers analogy questions.
+
+    \item \cite{Shazeer:MQA:2019}:
+
+    The paper introduces multi-query attention (MQA) transformer architecture.
+Multi-query attention is identical to multi-head attention except that different heads share the same set of keys and values.
+The idea behind this change to transformer architecture is to improve memory to computation ratio.
+The authors show that execution performance gain are noticeable while model quality remains about the same.
 \end{itemize}

### 2025-02-20 paper-reviews/papers-2025.tex
Add titans paper

\input{papers.tex}

\begin{document}

A total of \total{citenum} papers read in the year.

\section*{Machine Learning}
\begin{itemize}
    \item \cite{Behrouz:Titans:2024}:

    The authors present Titans --- a new machine learning model architecture that incorporates long-term memory with forgetting in order to be able to handle very large context windows. A forgetting mechanism is added to the model to avoid memory overflows. Long-term memory is added as a context to the current input sequence.
\end{itemize}

\end{document}

### 2025-02-21 paper-reviews/papers-2025.tex
Add DeepSeek-V3 paper

@@
 \begin{itemize}
+    \item \cite{DeepSeekAI:2025:DeepSeekV3}:
+
+    The report presents DeepSeek-V3 model architecture.
+The key difference from the Transformer architecture are: using Multi-Head Latent Attention (MLA) instead of MHA; using Mixture of Experts (MoE) instead of Feed Forward Network.
+DeepSeek-V3 MoE consists of shared and routed experts.
+Every token will run through the shared experts and select a few routed once.
+Each token may potentially end up using completely different routed experts.
+MLA implementation is aimed to reduce KV-cache usage by having a single shared cache for all heads.
+It may sound similar to Grouped Query Attention (GQA) but instead of compressing information from multiple heads into one, MLA moves per-head key and value information to per-head query thus only having the common for all heads information in keys and values.
+The authors also introduce Multi-Token Prediction (MTP) module that is used for training in the original work but can potentially be used for speculatively predicting several tokens in one run of the model.
+
     \item \cite{Behrouz:Titans:2024}:

### 2025-03-07 paper-reviews/papers-2025.tex
Add Mamba reference

@@
 Long-term memory is added as a context to the current input sequence.
+
+    \item \cite{Gu:Mamba:2024}:
+
+    The article presents Mamba model architecture that is aimed to address Transformer's computational inefficiency on long sequence lengths.
+The authors claim that the key weakness of the existing subquadratic-time architectures is inability to perform context-based reasoning.
+The mentioned problem is addressed by letting the structured state space model (SSM) parameters be functions of the input thus allowing the model to selectively propagate and forget information along the sequence length based on the current token.
+Mamba is based on these selective SSMs and requires neither attention nor MLP blocks.
+As results Mamba inference can scale linearly with sequence length.
+The key ability of Mamba is selectivity --- context-aware way to focus on or filter out inputs info sequential state.
+The authors achieve efficient utilization of the existing hardware through standard techniques: kernel fusion, parallel scan and recomputation.
 \end{itemize}

### 2025-03-07 paper-reviews/papers-2025.tex
Add pathologies of big data paper

@@
 
+\section*{Performance}
+\begin{itemize}
+    \item \cite{Jacobs:DataBasePerformance:2009}:
+
+    The article highlights optimization challenges that database developers face while working on big data analysis.
+The author demonstrates performance drawbacks of reading large databases for different kind of analyses.
+The article is largely focused on efficient usage of hardware capabilities: random vs.
+sequential access to storage devices.
+There is a good diagram that compares random and sequential accesses of hard drives, SSDs and DRAM.
+The author also shows few possible database architecture focusing on possible read requests that are supposed to be serviced by the database.
+There is also a good definition of "big data" in the very end of the paper.
+\end{itemize}
+
 \end{document}

### 2025-03-14 paper-reviews/papers-2025.tex
Add LoRE paper

@@
 The authors achieve efficient utilization of the existing hardware through standard techniques: kernel fusion, parallel scan and recomputation.
+
+    \item \cite{Hu:LoRA:2021}:
+
+    The article presents LoRA (Low-Rank Adaptation) technique for large language models.
+The technique is aimed to speedup fine-tuning and inference of the fine-tuned model.
+Traditionally, fine-tuning is used to update weights of a model pre-trained on a large set of generic data in order to specialize the model to handle a specific task.
+Basically it is a retraining of a generic model on a dataset specific for a desired task.
+With large language model it poses an obvious issue --- the fine-tuned model has the same number of parameters as the original model which makes the retraining task expensive and long.
+The main idea of LoRA is to freeze the pre-trained model weights and inject additional trainable rank decomposition matrices into the model in order to greatly reduce the number of trainable parameters.
+This approach allows easy switch of the model in production environment --- only the smaller additional matrices need to be switched.
+It is also possible to merge the trainable matrices into the original pre-trained model weights to avoid introducing extra latency during inference.
+The LoRA matrices consist of down projection from the original dimension to a small LoRA rank (the authors showed that even rank of two can be efficient) and up projection back to the original dimension.
+The high level idea is that the LoRA rank will record information specific for a desired task.
+The paper shows that it is preferable to adopt more weight matrices than to adopt a single weight matrix with a larger rank.
 \end{itemize}

### 2025-03-24 paper-reviews/papers-2025.tex
Add MLPerf Training paper

@@
 The paper shows that it is preferable to adopt more weight matrices than to adopt a single weight matrix with a larger rank.
+
+    \item \cite{Mattson:MLPerfTraining:2020}:
+
+    The paper present MLPerf training benchmark and describes challenges that are rather unique for machine learning workloads: stochastic nature of the training task; necessity to train to the completion to be able to compare model quality.
+For example, lower data precision from FP32 to BF16 will definitely improve training throughput but it may require a lot more iterations to complete the training task (achieve similar model quality).
+The benchmark includes a set of model architectures aimed to solve the most common machine learning tasks including image classification, object detection, translation (both transformer and RNN), etc.
 \end{itemize}

### 2025-04-01 paper-reviews/papers-2025.tex
Add sequence parallel paper

@@
 The benchmark includes a set of model architectures aimed to solve the most common machine learning tasks including image classification, object detection, translation (both transformer and RNN), etc.
+
+    \item \cite{Korthikanti:SequenceParallel:2022}:
+
+    The paper presents sequence parallel and selective activation recomputation techniques.
+The techniques are aimed to accelerate training of large Transformer models by reducing activation recomputation.
+The sequence parallel technique extends Megatron-LM idea by running previously sequential parts in sequence parallel manner.
+Basically, the Transformer layer is split into four interchanging tensor parallel and sequence parallel block.
+GEMMs run using tensor parallel while layer norm and dropout run using sequence parallel.
+Interesting to note that sequence parallel does not add any extra communication overhead because communication required for tensor parallel already does all operations needed for sequence parallel part.
+Ring all-reduce (required for tensor parallel) is composed of reduce-scatter and all-gather (both required for sequence parallel).
+One downside is that execution of reduce-scatter and all-gather combined is slower than all-reduce alone.
+In general, the paper can be used as a great source to get an idea of different model parallelization techniques.
 \end{itemize}

### 2025-04-02 paper-reviews/papers-2025.tex
Add KV-cache drop sparse attention paper

@@
 In general, the paper can be used as a great source to get an idea of different model parallelization techniques.
+
+    \item \cite{Wang:KvCacheDrop:2025}:
+
+    The paper describes an idea to optimize KV-cache size and attention calculations performed during decode (token generation) stage.
+The compressed version of KV-cache created during prefill (cache generation) stage consists of K and V results for fixed number of initial (first in input sequence) tokens, fixed number of recent tokens, last token and top-k of all the tokens in between initial initial and recent.
+The general idea is that initial and recent token are important and top-k selection can keep all the other important information while dropping other.
+Only recent tokens and last token in the KV-cache are getting updated during decode stage.
+Recent tokens acting like a ring buffer.
+Initial tokens and top-k stay the same.
+Thus KV-cache size always stays the same.
 \end{itemize}

### 2025-04-14 paper-reviews/papers-2025.tex
Add RoPE scaling laws paper

@@
 Thus KV-cache size always stays the same.
+
+    \item \cite{Liu:ScalingLawsRoPE:2024}:
+
+    The paper describes how Rotary Position Embedding (RoPE) is used by Transformer models to discern positional information in the input sequence.
+The authors studied extrapolation capabilities of Large Language Models based on RoPE with focus on modifications to rotary base.
+The paper proposed "Scaling Laws of RoPE-based Extrapolation", a framework to describe relations between the extrapolation performance and rotary base values as well as tunning context length.
 \end{itemize}

### 2025-04-20 paper-reviews/papers-2025.tex
Add Intel hides behind ICC paper
Add effective attention paper

@@
 The paper proposed "Scaling Laws of RoPE-based Extrapolation", a framework to describe relations between the extrapolation performance and rotary base values as well as tunning context length.
+
+    \item \cite{Luong:AttentionNMT:2015}:
+
+    The authors present earlier attention-based models for machine translation tasks.
+The paper considers global and local attention approaches.
+The global approach attends to all words in the input sequence while the local one only looks at a subset of input words at a time.
+The article proposes and compares several mechanisms to decide what words in input sequence to attend to for the local attention approach.
+The resulting model yields a new state-of-the-art result in English to German translation task.
 \end{itemize}
@@
 There is also a good definition of "big data" in the very end of the paper.
+
+    \item \cite{Gwennap:IntelHidesBehindICC:2017}:
+
+    The article highlights how Intel uses ICC compiler and library optimizations to submit higher SPEC CPU scores.
+The author emphasizes that it is important to use GCC --- an open sources, harder to manipulate compiler --- to achieve apple to apple comparison of different vendors' processors.
 \end{itemize}

### 2025-04-25 paper-reviews/papers-2025.tex
Add AdaComp paper

@@
 The resulting model yields a new state-of-the-art result in English to German translation task.
+
+    \item \cite{Chen:AdaComp:2018}:
+
+    The paper presents Adaptive Residual Gradient Compression (AdaComp) scheme that is based on Local Selection (LS) of gradient residuals.
+The main difference from LS is that AdaComp automatically adjust compression rate depending local configuration that includes number of learners, neural network architecture, layer types, mini-batch sizes and other parameters.
+Thus AdaComp provides soft compression threshold (allowing to self-adjust the compression rate) compared to LS that has the threshold as a fix parameter.
+The authors managed to achieve $~200X$ compression rate for fully-connected and recurrent layers, and $~40X$ compression rate for convolutional layers.
 \end{itemize}

### 2025-04-29 paper-reviews/papers-2025.tex
Add sequence to sequence paper

@@
 The authors managed to achieve $~200X$ compression rate for fully-connected and recurrent layers, and $~40X$ compression rate for convolutional layers.
+
+    \item \cite{Sutskever:Seq2seq:2014}:
+
+    The paper presents LSTM-based (Long Short-Term Memory) sequence to sequence model for machine translation.
+The main idea is to first translate \textit{reversed} input sequence to an intermediate fixed dimensioned vector representation and then translate the acquired representation to the target language.
+The input sequence is reversed to get shorter distance between input and target words during training.
+The resulting model performed better than advanced Statistical Machine Translation (SMT) system.
 \end{itemize}

### 2025-05-09 paper-reviews/papers-2025.tex
Add meta DLRM paper

@@
 The resulting model performed better than advanced Statistical Machine Translation (SMT) system.
+
+    \item \cite{Mudigere:MetaDLRMTraining:2022}:
+
+    The paper present Neo (software) and ZionEX (hardware) systems co-designed for Deep Learning Recommendation Model (DLRM) training.
+The very interesting feature of the DLRMs is that the models use lots of data-intensive embedding operators compared to mostly compute-intensive operations in conventional Deep Neural Networks (DNNs).
+Existing software and hardware solutions are typically targeting DNNs.
+Such frameworks are optimized for compute-intense workloads and miss critical optimizations for data-intense embedding operators.
+The authors' solution includes 4D parallelism (table-wise, row-wise, column-wise and data), high-performance embedding computation through fused kernels and special hardware design to achieve better communication and memory access speed.
 \end{itemize}

### 2025-05-13 paper-reviews/papers-2025.tex
Add SnapKV paper

@@
 The authors' solution includes 4D parallelism (table-wise, row-wise, column-wise and data), high-performance embedding computation through fused kernels and special hardware design to achieve better communication and memory access speed.
+
+    \item \cite{Li:SnapKV:2024}:
+
+    The paper presents KV cache compression technique called SnapKV.
+The compression is achieved by selecting important KV positions for each attention head.
+The key differences from other techniques are: SnapKV uses an observation windows consisting of several last tokens instead of just one; SnapKV uses pooling instead naive top feature results for top-k selection.
+The technique is based on an observation that each attention head in a model consistently focuses on specific prompt attention features.
+SnapKV allows to KV cache memory usage and compute demands during decoding but it does not affect prefill.
 \end{itemize}

### 2025-05-21 paper-reviews/papers-2025.tex
Add star attention paper

@@
 SnapKV allows to KV cache memory usage and compute demands during decoding but it does not affect prefill.
+
+    \item \cite{Acharya:StarAttention:2025}:
+
+    The paper presents distributed block-sparse approximate attention algorithm called Star Attention.
+The algorithm is broken into two phases.
+In the first phase, each node calculates block-wise local attention with context window segmented into blocks and each block prefixed with an anchor (first context block).
+In the seconds phase, query and local attention results are collected and accumulated to a global attention approximation by one selected node.
+The authors demonstrate that Star Attention can achieve substantial speedups without significant model accuracy degradation.
+Block size variation allows for speed/accuracy balancing.
 \end{itemize}

### 2025-05-22 paper-reviews/papers-2025.tex
Add SpeedRacer attack paper

@@
 
+\section*{Software Attacks}
+
+\begin{itemize}
+    \item \cite{Kallenberg:SpeedRacer:2014}:
+
+    The paper demonstrates how race conditions in Intel flash protection mechanism can be used to overwrite UEFI firmware.
+The attack utilizes race condition between a malicious thread running supervisor mode and BIOS protection thread running in SMM.
+The software threads can run on different hyper threads.
+Even though Intel chipsets provide a mechanism to protect against the attacks, neither reference UEFI implementation not UEFI provided by OEMs use the protection mechanisms as of the state when the paper was written.
+\end{itemize}
+
 \section*{Performance}

### 2025-05-28 paper-reviews/papers-2025.tex
Add ADL paper

@@
 
+\section*{Simulators}
+\begin{itemize}
+    \item \cite{Mishra:ADL:2005}:
+
+    Overview and classification of Architecture Description Languages (ADLs).
+Splits ADLs into structural (synthesis and validation oriented), behavioral (compiler and simulation oriented) and mixed.
+The paper includes an overview of typical workflows based on ADL.
+Overall it is a good reference for students to show importance and usefulness of special-purpose languages.
+\end{itemize}
+
 \end{document}

### 2025-06-05 paper-reviews/papers-2025.tex
Add CPU emulator detection paper review

@@
 Even though Intel chipsets provide a mechanism to protect against the attacks, neither reference UEFI implementation not UEFI provided by OEMs use the protection mechanisms as of the state when the paper was written.
+
+    \item \cite{Paleari:EmulatorDetection:2009}:
+
+    The paper presents and automated approach for detecting differences in behavior amount real CPUs and emulators.
+The differences can be used by malicious software to prevent itself from being analyzed by anti-virus tools.
+\texttt{or \%bh, 0x4(\%ebx)} is used as an example demonstrating different behavior on real CPU vs.
+QEMU.
+The example is really broken or it utilizes some hardware bug present in old x86 CPUs.
+\texttt{\%bh} register is initialized to \texttt{0} prior the instruction execution thus making the instruction's behavior effectively equal to \texttt{nop} but the paper claims that it changes \texttt{0x4(\%ebx)} memory location on real CPU and doesn't change it in QEMU.
+The example C code corresponds to an invalid C program but can be easily modified to become valid.
+The modified program doesn't experience the behavior claimed in the paper when the program is executed on a real CPU (Intel(R) Xeon(R) Gold 6248R CPU).
+Modified program code:
+
+    \lstinputlisting[language=C]{}
 \end{itemize}

### 2025-06-05 paper-reviews/red_pill_example.c

@@ -1,0 +1,34 @@
+#include <stdlib.h>
+#include <string.h>
+#include <signal.h>
+#include <sys/mman.h>
+#include <assert.h>
+
+// Catch page-fault exceptions
+void pf_handler(int i) {
+    exit(1);
+}
+
+void main() {
+    // Initialize the state:
+    // -registers
+    asm volatile("mov $0x1, %eax");
+    asm volatile("mov $0x08090004, %ebx");
+    // - memory
+    size_t page_size = 4096;
+    void *page = mmap((void*)0x08090000, page_size, PROT_READ | PROT_WRITE, MAP_ANONYMOUS | MAP_PRIVATE, -1, 0);
+    assert(page);
+    memcpy((void*)0x08090008, "\x11\x22", 2);
+    // - setup handlers to catch exceptions
+    signal(SIGSEGV, pf_handler);
+    //Execute the code: or %bh, 0x4(%ebx)
+    asm volatile(".byte 0x08,0x7c,0xe3,0x04");
+    // Compare the state (only the potentially
+    // differing bytes) and exit accordingly
+    if (memcmp((void*)0x08090008,"\x11\x26", 2) == 0) {
+        munmap(page, page_size);
+        exit(0);
+    }
+    munmap(page, page_size);
+    exit(1);
+}

### 2025-07-04 paper-reviews/papers-2025.tex
Add AVX512-based Base64 paper

@@
 The author emphasizes that it is important to use GCC --- an open sources, harder to manipulate compiler --- to achieve apple to apple comparison of different vendors' processors.
+
+    \item \cite{Wojciech:Avx512Base64:2019}:
+
+    The paper presents AVX512-based base64 encoding and decoding implementation.
+The implementation achieves five and seven-fold improvement compared to AVX2-based implementation for decoding and encoding correspondingly.
+The implementation uses AVX512 BMI instructions in a nice way that allows to decode/encode base64 at speed close to memory copy.
 \end{itemize}

### 2025-07-18 paper-reviews/papers-2025.tex
Add approximate computing paper

@@
 The implementation uses AVX512 BMI instructions in a nice way that allows to decode/encode base64 at speed close to memory copy.
+
+    \item \cite{Agrawal:ApproximateComputing:2016}:
+
+    The paper explores approximate computing techniques applied to a set applications that do not require exact calculations.
+The techniques explored in the paper include: loop perforation (omitting executing of some loop iterations), reduced precision for floating point calculations and relaxed synchronization for concurrent execution.
+The authors show that the techniques can be combined together in many cases.
+How and what techniques are to be used depend on the application.
+The resulting performance improvements also varies with the application.
 \end{itemize}
